\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-02-11 doc/publications/plos16/plos16.tex
upgraded to new version of plot template

% Template for PLoS
% Version 3.3 June 2016
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example
% table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
%
% Below are a few tips to help format your equations and other special characters according to our specifications. For
% more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at
% http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example,
% x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as
% CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part
% of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change
% "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

@@

@@

%% Additional packages
\usepackage{footmisc}
\usepackage{multirow}
\usepackage[T1]{fontenc}
%% required the installation of cm-super 55MB - TeX font package (full version) with CM (EC) in Type1 in T1, T2*, TS1, X2 enc

%%local edits (rob)
\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%hv
{\raggedright\footnotesize #1}}
\reversemarginpar
\setlength{\marginparwidth}{5cm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{xspace} % see http://tex.stackexchange.com/questions/31091/space-after-latex-commands

@@

As shown in Figure~\ref{figure:ranksumtest}, the known genes are ranked significantly higher than variants from other locations in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$), supporting our hypothesis that biologically relevant features are contributing more information to our classification model.

The gene with the highest-ranked features is COLEC10 (collectin 10) as shown in Figure~\ref{figure:knowngeneranks}, a well established contributor to BDM~\cite{Kemp2014,Liu2008} with verified transcription factor binding site for FOXL1, also a known BDM gene~\cite{DECODE}.
The second highest ranked gene is ZBTB40 on 22q11. Stagi {\it et al.}~\cite{Stagi.2010}, observe that individuals with a
micro-deletion in this region have a significant reduction in bone mass. In this same region, we also observe the next highest ranked gene, PRODH (proline dehydrogenase). 
PRODH catalyses the first step in proline degradation, and as
proposed by Phang {\it et al.}~\cite{Phang.2008}, proline is made available by collagen degradation.

The lowest ranked gene with literature implied BMD association is DCDC5 (Doublecortin Domain Containing 5) ranked 9,830 out of 10,000 (see Figure~\ref{figure:knowngeneranks}). 
Similar to the conclusions drawn by Duncan {\it et al.}~\cite{Duncan.2011}, our results also indicate that this gene does not have a strong association with BMD. 
In fact, DCDC5 is not highly expressed in bone, and any mechanisms by which this gene might regulate BMD are unclear \cite{Thakker2012}. 
Although variants near DCDC5 are associated with BMD \cite{Rivadeneira2009}, it is questionable as to whether the gene itself contributes to BMD.

Hence focusing on the top 20 most associated variants, the majority are from five known BMD genes (left side of the vertical line in Figure~\ref{figure:knowngeneranks}), however there are also five novel locations as annotated in Figure~\ref{figure:ranksumtest}. Most notably the SNP in the intronic region of Unc13c (rs12916554) is linked to quality of sleep~\cite{Nicod2016}, which in turn is reported to be the best predictor of osteoporosis risk~\cite{Magali2016}, a condition diagnosed by measuring BMD.  
Also listed is rs1035665, which was identified in a recent GWAS study to regulate ABCF1 (ATP Binding Cassette Subfamily F Member 1)~\cite{Dixon2007} with member 2 of this subfamily being a known BMD gene. The other three novel loci are on or near ncRNA.

>>>>

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in,showframe]{geometry}

@@

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% Additional packages and local edits
\usepackage{footmisc}
\usepackage{multirow}
\usepackage[T1]{fontenc}
%% required the installation of cm-super 55MB - TeX font package (full version) with CM (EC) in Type1 in T1, T2*, TS1, X2 enc

\let\oldmarginpar\marginpar
\renewcommand\marginpar[1]{\-\oldmarginpar[\raggedleft\footnotesize #1]%hv
{\raggedright\footnotesize #1}}
\reversemarginpar

\setlength{\marginparwidth}{5cm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{xspace} % see http:
%\usepackage{showframe}

@@

@@

As shown in Fig~\ref{figure:ranksumtest}, the known genes are ranked significantly higher than variants from other
locations in the genome (Mann-Whitney U, $p=1.3\times10^{-07}$), supporting our hypothesis that biologically relevant
features are contributing more information to our classification model.

The gene with the highest-ranked features is COLEC10 (collectin 10) as shown in Fig~\ref{figure:knowngeneranks}, a
well established contributor to BDM~\cite{Kemp2014,Liu2008} with verified transcription factor binding site for FOXL1,
also a known BDM gene~\cite{DECODE}.  The second highest ranked gene is ZBTB40 on 22q11. Stagi {\it et
  al.}~\cite{Stagi.2010}, observe that individuals with a micro-deletion in this region have a significant reduction in
bone mass. In this same region, we also observe the next highest ranked gene, PRODH (proline dehydrogenase).  PRODH
catalyses the first step in proline degradation, and as proposed by Phang {\it et al.}~\cite{Phang.2008}, proline is
made available by collagen degradation.

The lowest ranked gene with literature implied BMD association is DCDC5 (Doublecortin Domain Containing 5) ranked 9,830
out of 10,000 (see Fig~\ref{figure:knowngeneranks}).  Similar to the conclusions drawn by Duncan {\it et
  al.}~\cite{Duncan.2011}, our results also indicate that this gene does not have a strong association with BMD.  In
fact, DCDC5 is not highly expressed in bone, and any mechanisms by which this gene might regulate BMD are unclear
\cite{Thakker2012}.  Although variants near DCDC5 are associated with BMD \cite{Rivadeneira2009}, it is questionable as
to whether the gene itself contributes to BMD.

Hence focusing on the top 20 most associated variants, the majority are from five known BMD genes (left side of the
vertical line in Fig~\ref{figure:knowngeneranks}), however there are also five novel locations as annotated in
Fig~\ref{figure:ranksumtest}. Most notably the SNP in the intronic region of Unc13c (rs12916554) is linked to quality
of sleep~\cite{Nicod2016}, which in turn is reported to be the best predictor of osteoporosis risk~\cite{Magali2016}, a
condition diagnosed by measuring BMD.  Also listed is rs1035665, which was identified in a recent GWAS study to regulate
ABCF1 (ATP Binding Cassette Subfamily F Member 1)~\cite{Dixon2007} with member 2 of this subfamily being a known BMD
gene. The other three novel loci are on or near ncRNA.
